 

### 2023-03-05 intro.Rmd
New figure for Introduction

At all levels of sociodemographic development, this burden of morbidity and mortality is rising.[@patelLancetCommissionGlobal2018] Yet despite this global scale and despite a growing evidence base for impactful, cost-effective interventions that may help,[@patelMentalNeurologicalSubstance2016] the profound impact of mental disorders has been met by a profound treatment gap.[@demyttenaerePrevalenceSeverityUnmet2004] The majority of people living with a mental disorder do not have access to adequate care,[@bassUnitedNationsGeneral2012a; @demyttenaerePrevalenceSeverityUnmet2004] and many face stigma,[@alonsoPerceivedStigmaIndividuals2009; @angermeyerStigmaMentalIllness2003] discrimination,[@stuartMentalIllnessEmployment2006; @thornicroftDiscriminationHealthCare2007] and abuse.[@humanrightswatchDeathSentenceAbuses2012; @mfoafo-mcarthyHumanRightsViolations2014; @rahmanMentalDistressHuman2021]

>>>>

At all levels of sociodemographic development, this burden of morbidity and mortality is rising.[@patelLancetCommissionGlobal2018] Yet despite this global scale and despite a growing evidence base for impactful, cost-effective interventions that may help,[@patelMentalNeurologicalSubstance2016] the profound impact of mental disorders has been met by a profound treatment gap.[@demyttenaerePrevalenceSeverityUnmet2004] The majority of people living with a mental disorder do not have access to adequate care (Figure \ref{figure_service_gap}),[@bassUnitedNationsGeneral2012a; @demyttenaerePrevalenceSeverityUnmet2004] and many face stigma,[@alonsoPerceivedStigmaIndividuals2009; @angermeyerStigmaMentalIllness2003] discrimination,[@stuartMentalIllnessEmployment2006; @thornicroftDiscriminationHealthCare2007] and abuse.[@humanrightswatchDeathSentenceAbuses2012; @mfoafo-mcarthyHumanRightsViolations2014; @rahmanMentalDistressHuman2021]

\blandscape

```{r figure_service_gap, include = TRUE, echo=FALSE, warnings = FALSE, fig.width = 11, fig.height = 6.5, fig.align='center', out.width = '97%',  fig.scap="Countries where fewer than 10 psychiastrists per 100,000 capita were working in the health sector, 2013-2017.", out.extra='',  fig.cap="\\label{figure_all_diseases}Countries where fewer than 10 psychiastrists per 100,000 capita were working in the health sector between 2013-2017, average. Data from the Globa Health Observator (GHO) of the World Health Organization (WHO)."}

# Loading libraries
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(sp)
library(broom)
library(sf)
library(scales)
library(viridis)
library(data.table)
library(ggpubr)
library(lemon)
library(rnaturalearth)
library(rnaturalearthdata)
library(sysfonts)
library(showtext)
font_add_google("Source Sans Pro")
library(gridExtra)
showtext_auto() 
showtext_opts(dpi = 300)  
library(rnaturalearthhires) # install.packages("rnaturalearthhires", repos = "http://packages.ropensci.org", type = "source")
library(ggtext)

world_map <- ne_countries(scale = "large", returnclass = "sf")

who_gho_data <- read.csv("C:/Users/danie/Dropbox (Harvard University)/Important Files/3. PhD/Thesis/3. Draft/1. RMarkdown/phd_thesis/who_gho_data.csv") %>% rename_with(~ gsub("_working_in_mental_health_sector_.per_100_000_population.", "", .), 
              contains("_working_in_mental_health_sector_.per_100_000_population.")) %>%
    pivot_longer(cols = -country, 
               names_to = c("type", "year"), 
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  group_by(country, type) %>%
  mutate(average = mean(value, na.rm = TRUE)) %>%
  ungroup() %>%
  select(average, country, type) %>%
  unique() %>%
  filter(type == "psychiatrists")

country <- c("Afghanistan", "Albania", "Angola", "Antigua and Barbuda", "Argentina", "Armenia", "Australia", "Azerbaijan", "Bahrain", "Bangladesh", "Barbados", "Belarus", "Belgium", "Belize", "Bhutan", "Bolivia (Plurinational State of)", "Bosnia and Herzegovina", "Brazil", "Brunei Darussalam", "Bulgaria", "Burkina Faso", "Burundi", "Cambodia", "Canada", "Central African Republic", "Chad", "Chile", "China", "Colombia", "Comoros", "Congo", "Cook Islands", "Costa Rica", "Côte d'Ivoire", "Croatia", "Cuba", "Cyprus", "Czechia", "Dominican Republic", "Ecuador", "Egypt", "El Salvador", "Equatorial Guinea", "Eritrea", "Estonia", "Eswatini", "Ethiopia", "Fiji", "Finland", "France", "Gabon", "Gambia", "Georgia", "Germany", "Ghana", "Greece", "Grenada", "Guatemala", "Guinea", "Guinea-Bissau", "Guyana", "Haiti", "Honduras", "Hungary", "India", "Indonesia", "Iran (Islamic Republic of)", "Iraq", "Israel", "Italy", "Jamaica", "Japan", "Jordan", "Kenya", "Kiribati", "Kyrgyzstan", "Latvia", "Lebanon", "Liberia", "Lithuania", "Luxembourg", "Madagascar", "Malaysia", "Maldives", "Mali", "Marshall Islands", "Mauritania", "Mauritius", "Mexico", "Monaco", "Mongolia", "Montenegro", "Morocco", "Mozambique", "Myanmar", "Namibia", "Nauru", "Nepal", "Netherlands", "New Zealand", "Nicaragua", "Niger", "Norway", "Oman", "Panama", "Peru", "Philippines", "Poland", "Qatar", "Republic of Korea", "Republic of Moldova", "North Macedonia", "Romania", "Russian Federation", "Rwanda", "Saint Kitts and Nevis", "Saint Lucia", "Saint Vincent and the Grenadines", "Samoa", "Sao Tome and Principe", "Saudi Arabia", "Senegal", "Serbia", "Seychelles", "Sierra Leone", "Singapore", "Slovenia", "South Africa", "South Sudan", "Spain", "Sri Lanka", "Sudan", "Suriname", "Sweden", "Switzerland", "Syrian Arab Republic", "Tajikistan", "Thailand", "Timor-Leste", "Togo", "Tonga", "Tunisia", "Türkiye", "Uganda", "Ukraine", "United Arab Emirates", "United Republic of Tanzania", "United States of America", "Uruguay", "Uzbekistan", "Vanuatu", "Yemen", "Zambia", "Zimbabwe")

adm0_a3 <- c("AFG", "ALB", "AGO", "ATG", "ARG", "ARM", "AUS", "AZE", "BHR", "BGD", "BRB", "BLR", "BEL", "BLZ", "BTN", "BOL", "BIH", "BRA", "BRN", "BGR", "BFA", "BDI", "KHM", "CAN", "CAF", "TCD", "CHL", "CHN", "COL", "COM", "COG", "COK", "CRI", "CIV", "HRV", "CUB", "CYP", "CZE", "DOM", "ECU", "EGY", "SLV", "GNQ", "ERI", "EST", "SWZ", "ETH", "FJI", "FIN", "FRA", "GAB", "GMB", "GEO", "DEU", "GHA", "GRC", "GRD", "GTM", "GIN", "GNB", "GUY", "HTI", "HND", "HUN", "IND", "IDN", "IRN", "IRQ", "ISR", "ITA", "JAM", "JPN", "JOR", "KEN", "KIR", "KGZ", "LVA", "LBN", "LBR", "LTU", "LUX", "MDG", "MYS", "MDV", "MLI", "MHL", "MRT", "MUS", "MEX", "MCO", "MNG", "MNE", "MAR", "MOZ", "MMR", "NAM", "NRU", "NPL", "NLD", "NZL", "NIC", "NER", "NOR", "OMN", "PAN", "PER", "PHL", "POL", "QAT", "KOR", "MDA", "MKD", "ROU", "RUS", "RWA", "KNA", "LCA", "VCT", "WSM", "STP", "SAU", "SEN", "SRB", "SYC", "SLE", "SGP", "SVN", "ZAF", "SSD", "ESP", "LKA", "SDN", "SUR", "SWE", "CHE", "SYR", "TJK", "THA", "TLS", "TGO", "TON", "TUN", "TUR", "UGA", "UKR", "ARE", "TZA", "USA", "URY", "UZB", "VUT", "YEM", "ZMB", "ZWE")

iso_crosswalk <- as.data.frame(cbind(country, adm0_a3))

who_gho_data <- who_gho_data %>%
  inner_join(., iso_crosswalk)

col_light_blue = "deeppink3"
col_grey = "#CCCCCC"
col_light_grey = "#F2F2F2"

world_map <- world_map %>%
  st_as_sf() %>%
  st_transform(CRS("+proj=robin"))

map <- world_map %>%
  filter(admin != "Antarctica") %>%
  full_join(who_gho_data, by = "adm0_a3") %>%
  mutate(fill = ifelse(average < 10,
                       "Fewer than 10 psychiatrists per 100,000 capita",
                       "Others")) %>%
  mutate(fill = ifelse(is.na(fill), "No data", fill)) %>%
  mutate(fill = factor(fill, 
                                   levels = c("Fewer than 10 psychiatrists per 100,000 capita", 
                                              "Others",
                                              "No data"))) %>%
  ggplot(aes(geometry = geometry, fill = fill)) +
  geom_sf(color = NA) +
  theme_void() +
  # labs(title = "Availability of psychiastrists",
  #      subtitle = " ") +
  scale_fill_manual(name = "",
                    values  = c(col_light_blue, col_grey, col_light_grey)) +
  theme(legend.position = "bottom",
        legend.justification='center',
        legend.direction='horizontal',
        text = element_text(family = "Arial"),
        # plot.title =element_text(size=18,   
        #                            face = "bold",
        #                            hjust = 0),
        axis.title.x = element_text(color=col_grey, size=10, hjust = 0),
        plot.caption = element_text(hjust = 0, 
                                          vjust = 1,
                                    face = "bold",
                                    size=10))

data <- world_map %>%
  filter(admin != "Antarctica") %>%
  full_join(who_gho_data, by = "adm0_a3") %>%
  mutate(category = ifelse(average < 10,
                           "A",
                           "B")) %>%
  select(category, pop_est) %>%
  st_drop_geometry() %>%
  group_by(category) %>%
  mutate(count = sum(as.numeric(pop_est), na.rm = TRUE)) %>%
  select(category, count) %>%
  ungroup() %>%
  unique()  %>%
  mutate(category = ifelse(is.na(category), "B", category)) %>%
  mutate(category = factor(category, 
                           levels = c("A", 
                                      "B"))) %>%
  arrange(category)

# Compute percentages
data$fraction <- data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax <- cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin <- c(0, head(data$ymax, n=-1))

# Make the plot
circle <- data %>%
  ggplot(aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
  geom_rect() +
  # geom_label( x=3.5, aes(y=labelPosition, label=label), size=6) +
  geom_richtext(aes(x = 0,
                    y = 0),
                label = paste0(round(data$fraction[data$category == "A"]*100,0),
                               "%"),
                color = "black",  
                fill = NA,
                label.color = NA,
                size = 10 / .pt,
                hjust = 0.5,
                vjust = .9) +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c(col_light_blue, col_grey))

text <- ggplot() +
  theme_void() +
  xlim(0, Inf) +
  geom_richtext(aes(x = 0,
                    y = 0),
                label = paste0('**',
                               "of the world's",
                               "**<br>**",
                               "population",
                               ",",
                               "**<br>",
                               paste0(round(data$count[data$category == "A"]/1000000000,1),
                                            " billion")),
                color = "black",  
                fill = NA,
                label.color = NA,
                size = 12 / .pt,
                hjust = 0,
                vjust = 0.5)

circle_text <- cowplot::plot_grid(circle, text, 
                     ncol = 2,
                     rel_heights =    c(7,1.3))

# title <- ggplot() +
#   geom_segment(aes(x = 0, y = 0, xend = .25, yend = 0), linewidth = .9, color = col_grey) +
#   # annotate("text", x = 0, y = 0, size=12, label = "Average, 2013-2017",
#   #          hjust = 0, vjust = -0.3,
#   #          color = "black", fontface = 'bold', family = "Martel") + 
#   theme_void() +
#   xlim(0, 0.25) +
#   ylim(-0.01, 0.04)

# combo <- cowplot::plot_grid(title, circle_text, 
#                             nrow = 2,
#                             rel_heights =  c(1.2,3)) 
#

# ggsave(map,
#        filename = paste0("gdp_",
#                          year,
#                          ".png"),
#        path = "C:/Users/danie/Dropbox (Harvard University)/Important Files/7. Data Science/mckinsey",
#        height = 7.5,
#        width= 12,
# #        bg = "white")
# 
# ggsave(combo,
#        filename = paste0("gdp_pie_",
#                          year,
#                          ".png"),
#        path = "C:/Users/danie/Dropbox (Harvard University)/Important Files/7. Data Science/mckinsey",
#        height = 2.05,
#        width= 2.25,
#        bg = "white")

final_figure <- cowplot::ggdraw(map) +
  cowplot::draw_plot(
    {circle_text}

    ,
    # The distance along a (0,1) x-axis to draw the left edge of the plot
    x = 0.05, 
    # The distance along a (0,1) y-axis to draw the bottom edge of the plot
    y = 0.20,
    # The width and height of the plot expressed as proportion of the entire ggdraw object
    width = 0.225, 
    height = 0.25)

print(final_figure)

```

\elandscape
